---
title: "Exercise-C5"
output: github_document
---
```{r}
rm(list=ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(rethinking)
library(dagitty)
library(dplyr)
```

# 5E1

$$
(2): \mu_{i} = \beta_{x}x_{i} + \beta_{z}z_{i}\\
(4): \mu_{i} = \alpha + \beta_{x}x_{i} + \beta_{z}z_{i}
$$

# 5E2

AD: Animal diversity
LA: Latitude
PD: Plant diversity

$$
AD_{i} = Func(LA_{i}, PD_{i})
$$

# 5E3

T: Time to PhD degree
F: Funding
S: Size of laboratory

$$
T_{i} = \alpha + \beta_{f}F_{i} + \beta_{s}S_{i} + \epsilon
$$

$$
\beta_{f} \geq 0\\
\beta_{s} \geq 0
$$

# 5E4

(1) + (3)

# 5M1

In the example, there is a spurious correlation between B and C. As A is the common cause of B and C. Regressing B on both A and C will show us the spurious correlation with a dramatic drop in the coefficient of C in model m5.1.3 in compared with model m5.1.2

```{r}
N <- 100
A <- rnorm(n=N, mean=0, sd=1)
B <- rnorm(n=N, mean=A, sd=1)
C <- rnorm(n=N, mean=A, sd=1)
```


```{r}
m5.1.1 <- quap(
  alist(
    B ~ dnorm(mu, sigma),
    mu <- a + b1*A,
    a ~ dnorm(0,1),
    b1 ~ dnorm(0,1),
    sigma ~ dexp(1)
  ),
  data = data.frame(B)
)
```


```{r}
precis(m5.1.1)
```

```{r}
m5.1.2 <- quap(
  alist(
    B ~ dnorm(mu, sigma),
    mu <- a + b1*C,
    a ~ dnorm(0,1),
    b1 ~ dnorm(0,1),
    sigma ~ dexp(1)
  ),
  data = data.frame(B)
)
```

```{r}
precis(m5.1.2)
```


```{r}
m5.1.3 <- quap(
  alist(
    B ~ dnorm(mu, sigma),
    mu <- a + b1*A + b2*C,
    a ~ dnorm(0,1),
    b1 ~ dnorm(0,1),
    b2 ~ dnorm(0,1),
    sigma ~ dexp(1)
  ),
  data = data.frame(B)
)
```


```{r}
precis(m5.1.3)
```

# 5M2

```{r}
N <- 100
G <- rnorm(n=N,mean=0,sd=1)
D <- rnorm(n=N,mean=G,sd=1)
E <- rnorm(n=N,mean=G,sd=1)
F <- rnorm(n=N,mean=D-E,sd=1)
```


```{r}
m5.2.1 <- quap(
  alist(
    F ~ dnorm(mu, sigma),
    mu <- a + b1*D,
    a ~ dnorm(0,1),
    b1 ~ dnorm(0,1),
    sigma ~ dexp(1)
  ),
  data = data.frame(F)
)
```


```{r}
precis(m5.2.1)
```


```{r}
m5.2.2 <- quap(
  alist(
    F ~ dnorm(mu, sigma),
    mu <- a + b1*E,
    a ~ dnorm(0,1),
    b1 ~ dnorm(0,1),
    sigma ~ dexp(1)
  ),
  data = data.frame(F)
)
```


```{r}
precis(m5.2.2)
```


```{r}
m5.2.3 <- quap(
  alist(
    F ~ dnorm(mu, sigma),
    mu <- a + b1*D + b2*E,
    a ~ dnorm(0,1),
    b1 ~ dnorm(0,1),
    b2 ~ dnorm(0,1),
    sigma ~ dexp(1)
  ),
  data = data.frame(F)
)
```


```{r}
precis(m5.2.3)
```


```{r}
m5.2.4 <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + b1*E,
    a ~ dnorm(0,1),
    b1 ~ dnorm(0,1),
    sigma ~ dexp(1)
  ),
  data = data.frame(D)
)
```


```{r}
precis(m5.2.4)
```
# 5M3

A high divorce rate implies more adults are available to match with each other. It possibly impacts the marriage rate.

However, it is hypothesized that the elders hesitate to get remarried because of the aging. Thus, we can regress the marriage rate on the divorce rate, conditional on the age.

# 5M4

```{r}
data("WaffleDivorce")
d <- WaffleDivorce
head(d)
```


```{r}
LDS.pop <- read.csv("/Users/mac/Documents/14. Bayes Statistic/Statistical Rethinking/data/LDS_population.csv")
```


```{r}
head(LDS.pop)
```


```{r}
for(i in d$Location){
  if(i %in% LDS.pop$State){}
  else{print(i)}
}
```

```{r}
rownames(LDS.pop)[rownames(LDS.pop) == "District Of Columbia"] <- "District of Columbia"
```

```{r}
LDS.pop[LDS.pop$State=="District Of Columbia","State"] <- "District of Columbia"
```


```{r}
for(i in d$Location){
  if(i %in% LDS.pop$State){}
  else{print(i)}
}
```


```{r}
final.d <- merge(d, LDS.pop, by.x="Location",by.y="State")
```

```{r}
colnames(final.d)
```

```{r}
write.csv(final.d, "LDS.divorce.csv")
```


```{r}
df <- final.d[,c("Divorce","MedianAgeMarriage","Marriage","Percent")]
head(df)
```


```{r}
df$std_divorce <- standardize(df$Divorce)
df$std_age <- standardize(df$MedianAgeMarriage)
df$std_marriage <- standardize(df$Marriage)
df$std_lds <- standardize(df$Percent)
```


```{r}
m5.4 <- quap(
  alist(
    std_divorce ~ dnorm(mean=mu, sd=sigma),
    mu <- a + bA*std_age + bM*std_marriage + bL*std_lds,
    a ~ dnorm(mean=0,sd=0.2),
    bA ~ dnorm(mean=0,sd=0.5),
    bM ~ dnorm(mean=0,sd=0.5),
    bL ~ dnorm(mean=0,sd=0.5),
    sigma ~ dexp(1)
  ),
  data=df
)
```


```{r}
precis(m5.4)
```

There is ambiguous impacts of marriage rates on divorce rates

# 5M5

P: Price
O: Obesity
E: Exercise
EO: Eat out

```{r}
dag5.5.1 <- dagitty("dag{P -> E; E -> O}")
coordinates(dag5.5.1) <- list(x=c(A=0,E=1,O=2),y=c(A=0,E=0,O=0))
drawdag(dag5.5.1)
```


```{r}
dag5.5.2 <- dagitty("dag{P -> EO; EO -> O}")
coordinates(dag5.5.2) <- list(x=c(A=0,EO=1,O=2),y=c(A=0,EO=0,O=0))
drawdag(dag5.5.2)
```

# 5H1 + 5H2

```{r}
h5.1 <- quap(
  alist(
    ## M -> A -> D
    std_divorce ~ dnorm(mu, sigma),
    mu <- a + bM*std_marriage + bA*std_age,
    a ~ dnorm(0,0.2),
    bM ~ dnorm(0,0.5),
    bA ~ dnorm(0,0.5),
    sigma ~ dexp(1),
    
    ## M -> A
    std_age ~ dnorm(mu_A, sigma_A),
    mu_A <- aA + bAM*std_marriage,
    aA ~ dnorm(0,0.2),
    bAM ~ dnorm(0,0.5),
    sigma_A ~ dexp(1)
    ),
  data=df
)
```


```{r}
precis(h5.1)
```
The conditional independency is that divorce rates are independent to marriage rates, conditionally in marriage ages. The data supports the statement.

# 5H3

```{r}
data(milk)
d <- milk
head(d)
```


```{r}
d$K <- standardize(d$kcal.per.g)
d$N <- standardize(d$neocortex.perc)
d$DM <- standardize(log(2*d$mass))
dcc <- d[complete.cases(d$K, d$N, d$DM),]
```


```{r}
h5.3 <- quap(
  alist(
    K ~ dnorm(mu, sigma),
    mu <- a + bDM*DM,
    a ~ dnorm(0,0.2),
    bDM ~ dnorm(0,0.5),
    sigma ~ dexp(1)
  ),
  data=dcc
)
```


```{r}
precis(h5.3)
```
Doubling M has no additional effect on K

# 5H4

```{r}
dag5h.4.1 <- dagitty("dag{A -> D; A -> M; M -> D; S -> M}")
coordinates(dag5h.4.1) <- list(x=c(A=0,M=1,D=0,S=1),y=c(A=0,M=0,D=1,S=1))
drawdag(dag5h.4.1)
```


```{r}
dag5h.4.2 <- dagitty("dag{A -> D; A -> M; M -> D; S -> M; S -> D}")
coordinates(dag5h.4.2) <- list(x=c(A=0,M=1,D=0,S=1),y=c(A=0,M=0,D=1,S=1))
drawdag(dag5h.4.2)
```


```{r}
dag5h.4.3 <- dagitty("dag{A -> D; A -> M; M -> D; S -> D}")
coordinates(dag5h.4.3) <- list(x=c(A=0,M=1,D=0,S=1),y=c(A=0,M=0,D=1,S=1))
drawdag(dag5h.4.3)
```


```{r}
df$State <- final.d$Location
head(df)
```


```{r}
Southern.states = c("Alabama","Arkansas","Delaware","Florida","Georgia","Kentucky","Louisiana","Maryland","Mississippi","North Carolina","Oklahoma","South Carolina","Tennessee","Texas","Virginia","West Virginia","District of Columbia")
```


```{r}
for(i in Southern.states){
  if(!(i %in% df$State)){
    print(i)
  }
}
```


```{r}
df$Southern <- as.factor(ifelse(df$State %in% Southern.states,2,1))
```


```{r}
head(df)
```


```{r}
h5.4 <- quap(
  alist(
    # D <- A -> M -> D <- S
    std_divorce ~ dnorm(mean=mu, sd=sigma),
    mu <- a[Southern] + bA*std_age + bM*std_marriage,
    a[Southern] ~ dnorm(mean=0,sd=0.2),
    bA ~ dnorm(mean=0,sd=0.5),
    bM ~ dnorm(mean=0,sd=0.5),
    sigma ~ dexp(1)
  ),
  data=df
)
```


```{r}
precis(h5.4, depth=2)
```

Divorce rates are not significantly different between Southern states and the rest

```{r}
```


```{r}
```


```{r}
```

